---
title: "02-names-values"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

# General

## Function factories

-   A **function factory** is a function that makes functions.
-   Function that are produced from a function factory are called **manufactured functions**.
    -   The key idea that makes them work is the enclosing environment of the manufactured function is an execution environment of the function factory.
-   We can see the environment using `rlang::env_print()` and extract values from the environment using `rlang::fn_env()`

```{r}
library(rlang)
power1 <- function(exp) {
  function(x) {
    x ^ exp
  }
}

square <- power1(2)
cube <- power1(3)

square
cube

square(3)
#> [1] 9
cube(3)
#> [1] 27

env_print(square)
#> <environment: 0x7fe851f7ccc0>
#> parent: <environment: global>
#> bindings:
#>  * exp: <dbl>

env_print(cube)
#> <environment: 0x7fe85508c390>
#> parent: <environment: global>
#> bindings:
#>  * exp: <dbl>

fn_env(square)$exp
#> [1] 2

fn_env(cube)$exp
#> [1] 3
```

### Forcing evaluation

```{r}
x <- 2
square <- power1(x)
x <- 3
square(2)
```

-   `force()` can be used to force the evaluation of an argument.

```{r}
power2 <- function(exp) {
  force(exp)
  function(x) {
    x ^ exp
  }
}

x <- 2
square <- power2(x)
x <- 3
square(2)
#> [1] 4
```

-   *Whenever you create a function factory, make sure every argument is evaluated, using force() as necessary if the argument is only used by the manufactured function.*

## Stateful functions

-   Function factories allow you to maintain state across function invocations.
-  This is possible because the environment of the manufactured function is unique and constant AND R has a **super assignment operator** (`<<-`) which modifies bindings in the enclosing environment.
- This operator rebinds an existing name found in a parent environment.

```{r}
new_counter <- function() {
  i <- 0
  
  function() {
    i <<- i + 1
    i
  }
}

counter_one <- new_counter()
counter_two <- new_counter()
counter_one()
#> [1] 1
counter_one()
#> [1] 2
counter_two()
#> [1] 1
```

- Stateful functions are best used in moderation. As soon as your function starts managing the state of multiple variables, it’s better to switch to R6.

# Concepts

# Useful functions

```{r}

```

# Exercises
