---
title: "02-names-values"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

# General

## Function factories

- A **function factory** is a function that makes functions.
- Function that are produced from a function factory are called **manufactured functions**.
  - The key idea that makes them work is the enclosing environment of the manufactured function is an execution environment of the function factory.
- We can see the environment using `rlang::env_print()` and extract values from the environment using `rlang::fn_env()`

```{r}
power1 <- function(exp) {
  function(x) {
    x ^ exp
  }
}

square <- power1(2)
cube <- power1(3)

square
cube

square(3)
#> [1] 9
cube(3)
#> [1] 27

env_print(square)
#> <environment: 0x7fe851f7ccc0>
#> parent: <environment: global>
#> bindings:
#>  * exp: <dbl>

env_print(cube)
#> <environment: 0x7fe85508c390>
#> parent: <environment: global>
#> bindings:
#>  * exp: <dbl>

fn_env(square)$exp
#> [1] 2

fn_env(cube)$exp
#> [1] 3
```



# Concepts

# Useful functions

```{r}

```

# Exercises
