---
title: "11-function-operators"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

# General

## Function operators

- **Function operators** take one or more functions as input and return a new function. Decorators is just another name for function operators which is what they are called in Python.
  - They can be thought of as just a function factory that takes a function as input. 
  - They're typically paired with functionals. 

```{r}
chatty <- function(f) {
  force(f)
  
  function(x, ...) {
    res <- f(x, ...)
    cat("Processing ", x, "\n", sep = "")
    res
  }
}
f <- function(x) x ^ 2
s <- c(3, 2, 1)

purrr::map_dbl(s, chatty(f))
#> Processing 3
#> Processing 2
#> Processing 1
#> [1] 9 4 1
```

## Existing function operators

- `purrr::safely()` provides a way to access results and errors in a function.
  - `safely()` always returns a list with two elements, result and error. If the function runs successfully, error is NULL and result contains the result; if the function fails, result is NULL and error contains the error.

```{r}
x <- list(
  c(0.512, 0.165, 0.717),
  c(0.064, 0.781, 0.427),
  c(0.890, 0.785, 0.495),
  "oops"
)

out <- rep(NA_real_, length(x))
for (i in seq_along(x)) {
  out[[i]] <- sum(x[[i]])
}
safe_sum <- safely(sum)
safe_sum
#> function (...) 
#> capture_error(.f(...), otherwise, quiet)
#> <bytecode: 0x7fafd9e2de58>
#> <environment: 0x7fafd9e2d9c0>
str(safe_sum(x[[1]]))
#> List of 2
#>  $ result: num 1.39
#>  $ error : NULL
str(safe_sum(x[[4]]))
#> List of 2
#>  $ result: NULL
#>  $ error :List of 2
#>   ..$ message: chr "invalid 'type' (character) of argument"
#>   ..$ call   : language .Primitive("sum")(..., na.rm = na.rm)
#>   ..- attr(*, "class")= chr [1:3] "simpleError" "error" "condition"
```

### Other purrr function operators

- `auto_browser()`: automatically executes browser() inside the function when there’s an error.

## Memoise cached computations

- `memoise::memoise()` memoises a function, meaning that the function will remember previous inputs and return cached results. This means it runs much faster but uses more memory.
- A relatively realistic use of memoisation is computing the Fibonacci series. This is because future calls can rely on previous computations

```{r}
fib <- function(n) {
  if (n < 2) return(1)
  fib(n - 2) + fib(n - 1)
}
system.time(fib(23))
#>    user  system elapsed 
#>   0.049   0.002   0.050
system.time(fib(24))
#>    user  system elapsed 
#>   0.069   0.002   0.070
fib2 <- memoise::memoise(function(n) {
  if (n < 2) return(1)
  fib2(n - 2) + fib2(n - 1)
})
system.time(fib2(23))
#>    user  system elapsed 
#>   0.009   0.000   0.008
```

## Interesting use case to download files

```{r eval = FALSE}
delay_by <- function(f, amount) {
  force(f)
  force(amount)
  
  function(...) {
    Sys.sleep(amount)
    f(...)
  }
}
system.time(runif(100))
#>    user  system elapsed 
#>       0       0       0
system.time(delay_by(runif, 0.1)(100))
#>    user  system elapsed 
#>    0.00    0.00    0.13

dot_every <- function(f, n) {
  force(f)
  force(n)
  
  i <- 0
  function(...) {
    i <<- i + 1
    if (i %% n == 0) cat(".")
    f(...)
  }
}
walk(1:100, runif)
walk(1:100, dot_every(runif, 10))
#> ..........

urls <- c(
  "adv-r" = "https://adv-r.hadley.nz", 
  "r4ds" = "http://r4ds.had.co.nz/"
  # and many many more
)
path <- paste(tempdir(), names(urls), ".html")

walk2(
  urls, path, 
  download.file %>% dot_every(10) %>% delay_by(0.1), 
  quiet = TRUE
)
```

# Concepts

# Useful functions

```{r}

```

# Exercises
