---
title: "14-R7"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

# General

## R6 101
- Methods belong to objects not generics. They are called like `object$method()`.
- R6 objects are mutable, meaning they're modified in place (not copied on modify) and have reference semantics.

## Using R6
- R6 only needs a single function call to create both the class and its methods. You use `R6::R6Class()` for this. 
  - The first argument is classname and must have UpperCamelCase names.
  - The second argument is public. It supplies a list of methods (functions) and fields (anything else) that make up the public interface of the object. Methods and fields must use snake_case. Methods can access the methods and fields of the current object via `self$`.
- You construct a new object from the class by calling the `new()` method
- Side-effect R6 methods should always return self invisibly. This allows us to chain together multiple method calls. This technique is called **method chaining** and is commonly used in languages like Python and JavaScript.

```{r}
library(R6)
Accumulator <- R6Class("Accumulator", list(
  sum = 0,
  add = function(x = 1) {
    self$sum <- self$sum + x 
    invisible(self)
  })
)
x <- Accumulator$new() 
x$sum
x$add(4) 
x$sum
#> [1] 4
x$add(5)$add(10)
x$sum
#> [1] 19
```

### Other important methods

- `$initialize()` overrides the default behaviour of `$new()`. It helps us to define rules for initializing values in the object.

```{r}
Person <- R6Class("Person", list(
  name = NULL,
  age = NA,
  initialize = function(name, age = NA) {
    stopifnot(is.character(name), length(name) == 1)
    stopifnot(is.numeric(age), length(age) == 1)
    
    self$name <- name
    self$age <- age
  }
))

hadley <- Person$new("Hadley", age = "thirty-eight")
#> Error in initialize(...): is.numeric(age) is not TRUE

hadley <- Person$new("Hadley", age = 38)
```

- More expensive validation requirements are implemented in a separate `$validate()` and only called when needed.

- `$print()` allows you to override the default printing behaviour. As with any R6 method called for its side effects, `$print()` should return `invisible(self)`.

```{r}
Person <- R6Class("Person", list(
  name = NULL,
  age = NA,
  initialize = function(name, age = NA) {
    self$name <- name
    self$age <- age
  },
  print = function(...) {
    cat("Person: \n")
    cat("  Name: ", self$name, "\n", sep = "")
    cat("  Age:  ", self$age, "\n", sep = "")
    invisible(self)
  }
))

hadley2 <- Person$new("Hadley")
hadley2
#> Person: 
#>   Name: Hadley
#>   Age:  NA
```

### Adding methods after creation

- It’s also possible to modify the fields and methods of an existing class.
- Add new elements to an existing class with `$set()`, supplying the visibility, the name, and the component
- New methods and fields are only available to new objects. They are not retrospectively added to existing objects.

```{r}
Accumulator <- R6Class("Accumulator")
Accumulator$set("public", "sum", 0)
Accumulator$set("public", "add", function(x = 1) {
  self$sum <- self$sum + x 
  invisible(self)
})
```

### Inheritance

- You can inherit behaviour from an existing class by providing the class object to the `inherit` argument.

```{r}
AccumulatorChatty <- R6Class("AccumulatorChatty", 
  inherit = Accumulator,
  public = list(
    add = function(x = 1) {
      cat("Adding ", x, "\n", sep = "")
      super$add(x = x)
    }
  )
)

x2 <- AccumulatorChatty$new()
x2$add(10)$add(1)$sum
#> Adding 10
#> Adding 1
#> [1] 11
```


### Privacy

- **Private** fields and methods are elements that can only be accessed from within the class, not from the outside.
- The `private` argument to R6Class works in the same way as the public argument.
- Fields and methods in private are available within the methods using `private$` instead of `self$`. However, you can't access them outside of the class.

```{r}
Person <- R6Class("Person", 
  public = list(
    initialize = function(name, age = NA) {
      private$name <- name
      private$age <- age
    },
    print = function(...) {
      cat("Person: \n")
      cat("  Name: ", private$name, "\n", sep = "")
      cat("  Age:  ", private$age, "\n", sep = "")
    }
  ),
  private = list(
    age = NA,
    name = NULL
  )
)

hadley3 <- Person$new("Hadley")
hadley3
#> Person: 
#>   Name: Hadley
#>   Age:  NA
hadley3$name
#> NULL
```


### Active fields

- **Active fields** allow you to define components that look like fields from the outside. However, they are defined with functions, like methods.

```{r}
Rando <- R6::R6Class("Rando", active = list(
  random = function(value) {
    if (missing(value)) {
      runif(1)  
    } else {
      stop("Can't set `$random`", call. = FALSE)
    }
  }
))
x <- Rando$new()
x$random
#> [1] 0.0808
x$random
#> [1] 0.834
x$random
#> [1] 0.601
x$random <- 5
#> Error: Can't set `$random`
```


### Reference semantics

- Reference semantics means that R6 objects are not copied when modified.
- We need to use `$clone()` if we want a copy.

```{r}
y1 <- Accumulator$new() 
y2 <- y1

y1$add(10)
c(y1 = y1$sum, y2 = y2$sum)
#> y1 y2 
#> 10 10
y2 <- y1$clone()

y1$add(10)
c(y1 = y1$sum, y2 = y2$sum)
#> y1 y2 
#> 10  0
```

### Finalizers

- Finalizers usually play a similar role to `on.exit()`, cleaning up any resources created by the initializer.
- It will be run when the object is deleted (or more precisely, by the first garbage collection after the object has been unbound from all names) or when R exits.

```{r}
TemporaryFile <- R6Class("TemporaryFile", list(
  path = NULL,
  initialize = function() {
    self$path <- tempfile()
  },
  finalize = function() {
    message("Cleaning up ", self$path)
    unlink(self$path)
  }
))
```

```{r}
tf <- TemporaryFile$new()
rm(tf)
#> Cleaning up /tmp/Rtmpk73JdI/file155f31d8424bd
```



# Concepts

# Useful functions

```{r}

```

# Exercises
