---
title: "20-evaluation"
format: html
editor_options: 
  chunk_output_type: console
---

# General

## Evaluation intro

- `eval()` evaluates an expression in an environment
- a quosure is a data structure that captures an expression along with its associated environment, as found in function arguments
- a data mask makes it easier to evaluate an expression in the context of a data frame
- quasiquotation, quosures, and data masks form what we call **tidy evaluation**

## How `eval()` works

- `eval()` has two arguments `expr` and `envir`:
  - the object to evaluate, typically a symbol or expression
    - this first argument is evaluated, not quoted
  - and the environment in which the expression should be evaluated
    - by default, this is the current environment, i.e. the calling environment
  
## Using `local()`

- Intermediate variables have no long-term use and could be quite large, so you’d rather not keep them around. 
- One approach is to clean up after yourself using rm(); another is to wrap the code in a function and just call it once. A more elegant approach is to use local().

```{r}
# Clean up variables created earlier
rm(x, y)

foo <- local({
  x <- 10
  y <- 200
  x + y
})

foo
#> [1] 210
x
#> Error in eval(expr, envir, enclos): object 'x' not found
y
#> Error in eval(expr, envir, enclos): object 'y' not found
```

  
## Quosures

- there isn't a data structure that can hold both an expression and an environment
- rlang fills the gap with the quosure, an object that contains an expression and an environment
  - the name is a portmanteau of quoting and closure, because a quosure both quotes the expression and encloses the environment
- three ways to create quosures:  `enquo()`, `quo()`, and `new_quosure()`

```{r}
## enquo() and enquos() to capture user-supplied expressions
## vast majority of quosures should be created this way
foo <- function(x) enquo(x)
foo(a + b)
#> <quosure>
#> expr: ^a + b
#> env:  global
## quo() and quos() exist to match to expr() and exprs() 
## used very rarely
quo(x + y + z)
#> <quosure>
#> expr: ^x + y + z
#> env:  global
## new_quosure() create a quosure from its components: an expression and an environment
## also rarely needed in practice
new_quosure(expr(x + y), env(x = 1, y = 10))
#> <quosure>
#> expr: ^x + y
#> env:  0x7fac62d44870
```

## Evaluating quosures

- Quosures are paired with a new evaluation function `eval_tidy()` that takes a single quosure instead of an expression-environment pair

```{r}
q1 <- new_quosure(expr(x + y), env(x = 1, y = 10))
eval_tidy(q1)
#> [1] 11
```

## Data masks

- Data mask is a data frame where the evaluated code will look first for variable definitions
- It allows you to mingle variables from an environment and a data frame in a single expression

```{r}
q1 <- new_quosure(expr(x * y), env(x = 100))
df <- data.frame(y = 1:10)

eval_tidy(q1, df)
#>  [1]  100  200  300  400  500  600  700  800  900 1000
#>  
```

## Data mask pronouns

- data mask provides two pronouns: .data and .env.:
  - .data$x always refers to x in the data mask.
  - .env$x always refers to x in the environment


# Concepts

# Useful functions

```{r}

```

# Exercises
