---
title: "03-subsetting"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

# Useful functions

```{r}

```

# Exercises

What is the result of subsetting a vector with positive integers, negative integers, a logical vector, or a character vector?

What’s the difference between [, [[, and $ when applied to a list?

When should you use drop = FALSE?

If x is a matrix, what does x[] <- 0 do? How is it different from x <- 0?

How can you use a named vector to relabel categorical variables?

Fix each of the following common data frame subsetting errors:


```{r}
mtcars[mtcars$cyl = 4, ]
mtcars[-1:4, ]
mtcars[mtcars$cyl <= 5]
mtcars[mtcars$cyl == 4 | 6, ]
```

Why does the following code yield five missing values? (Hint: why is it different from x[NA_real_]?)
```{r}
x <- 1:5
x[NA]
#> [1] NA NA NA NA NA
```

What does upper.tri() return? How does subsetting a matrix with it work? Do we need any additional subsetting rules to describe its behaviour?

```{r}
x <- outer(1:5, 1:5, FUN = "*")
x[upper.tri(x)]
```

Why does mtcars[1:20] return an error? How does it differ from the similar mtcars[1:20, ]?

Implement your own function that extracts the diagonal entries from a matrix (it should behave like diag(x) where x is a matrix).

What does df[is.na(df)] <- 0 do? How does it work?

Brainstorm as many ways as possible to extract the third value from the cyl variable in the mtcars dataset.

Given a linear model, e.g., mod <- lm(mpg ~ wt, data = mtcars), extract the residual degrees of freedom. Then extract the R squared from the model summary (summary(mod))

How would you randomly permute the columns of a data frame? (This is an important technique in random forests.) Can you simultaneously permute the rows and columns in one step?

How would you select a random sample of m rows from a data frame? What if the sample had to be contiguous (i.e., with an initial row, a final row, and every row in between)?

How could you put the columns in a data frame in alphabetical order?

# Playground
